Bayesian Sparsification Methods for Deep Complex-valued Networks

### 2020-06-23 paper/paper.tex
10 added acknowledgements and software sections
edited appendix-C

The source code for the experiments and the figures in this study is available at
\url{https://github.com/ivannz/complex_paper/tree/v2020.6}.
%
The source code for a package based on PyTorch \citep{paszke_pytorch_2019}, which
implements $\cplx$-valued Sparse Variational Dropout and ARD layers and provides other
basic layers for $\cplx$VNN is available at
\url{https://github.com/ivannz/cplxmodule}.
\smallskip

@@

@@

%notation
By $e_i$ we denote the $i$-th real unit vector of dimensionality \emph{conforming} to the
matrix-vector expression it is used in. Let $\vec{M}$ denote \emph{row-major} flattening
of $M$ into a vector, i.e. in lexicographic order of its indices. $\diag{(\cdot)}$ embeds
$\cplx^n$ into $n\times n$ diagonal matrices, and $\otimes$ denotes the Kronecker product,
for which we note the following identities $
  \vec{P Q R} = (P \otimes R^\top) \vec{Q}
$, and $
  (P \otimes R) (C \otimes S) = P Q \otimes R S
$ \citep{petersen_matrix_2012}.
% It follows from the definition of the Kronecker product
% of $P$ and $R^\top$ and the row-major order vectorization.

If we assume a mean field $\cplx$-Gaussian approximate posterior for $
  W \in \cplx^{n\times m}
$, then
\begin{equation}  \label{eq:c-gauss-vi-general-vec}
  \vec{W}
    \sim \mathcal{C}\mathcal{N}_{nm} \bigl(
      \vec{\mu}, \diag{\vec{\Sigma}}, \diag{\vec{C}}
    \bigr)
  \,,
\end{equation}
where mean, \emph{variance} and relation are given by matrices $
  \mu, \Sigma, C \in \cplx^{n\times m}
$ with $\Sigma_{ij} \geq 0$, and $
  \lvert C_{ij} \rvert^2 \leq \Sigma_{ij}
$. For any $x \in \cplx^m$ and $b\in \cplx^n$ $
  y = W x + b
    = (I_n \otimes x^\top) \vec{W} + b
$, whence the covariance and relation matrices of $y$ are
\begin{align}
  (I \otimes x^\top) \diag{\vec{\Sigma}} \conj{(I \otimes x^\top)^\top}
    % = \sum_{ij} (I \otimes x^\top) e_i \otimes e_j
    %   \Sigma_{ij} (e_i \otimes e_j)^\top (I \otimes x^\top)^{\hop}
    % = \sum_{ij} (e_i \otimes x^\top e_j)
    %   \Sigma_{ij} (e_i \otimes x^\top e_j)^{\hop}
    % = \sum_{ij} (e_i \otimes x_j) \Sigma_{ij} (e_i^\top \otimes \conj{x}_j)
    & = \sum_{i=1}^n (e_i e_i^\top) \sum_{j=1}^m \Sigma_{ij} x_j \conj{x}_j
    \,,  \label{eq:cn-gauss-lrt-cov} \\
  (I \otimes x^\top) \diag{\vec{C}} (I \otimes x^\top)^\top
    % = \sum_{ij} (e_i \otimes x_j) C_{ij} (e_i \otimes x_j)^\top
    & = \sum_{i=1}^n (e_i e_i^\top) \sum_{j=1}^m C_{ij} x_j^2
    \,.  \label{eq:cn-gauss-lrt-rel}
\end{align}
Since \eqref{eq:cn-gauss-lrt-cov} and \eqref{eq:cn-gauss-lrt-rel} are diagonal, $(y_i)_{i=1}^n$
are independent whence \eqref{eq:cn-affine} implies
\begin{equation}  \label{eq:cplx-gauss-trick-appendix}
  y_i
    \sim \mathcal{CN}
      \bigl(
        e_i^\top \mu x + b_i,
        \, \sum_{j=1}^m \Sigma_{ij} \lvert x_j \rvert^2,
        \, \sum_{j=1}^m C_{ij} x_j^2
      \bigr)
    \,.
\end{equation}
Hence \eqref{eq:cplx-gauss-trick} follows.

>>>>

@@

\section*{Software and Data} % (fold)
\label{sec:software_and_data}

% we strongly encourage the publication of software and data with the camera-ready version
% of the paper whenever appropriate. This can be done by including a URL in the copy.
% do not include URLs that reveal your institution or identity in your submission for review
% * provide an anonymous URL or upload the material as ``Supplementary Material'' into the CMT

The source code for a package based on PyTorch \citep{paszke_pytorch_2019}, which
implements $\cplx$-valued Sparse Variational Dropout and ARD layers and provides other
basic layers for $\cplx$VNN is available at
\url{https://github.com/ivannz/cplxmodule}.
%
The source code for the experiments and the figures in this study is available at
\url{https://github.com/ivannz/complex_paper/tree/v2020.6}.

% section* software_and_data (end)

% Acknowledgements should only appear in the accepted version.
\section*{Acknowledgements} % (fold)
\label{sec:acknowledgements}

% Do not include acknowledgements in the initial version of the paper submitted for blind review.

% the final camera-ready version can (and probably should) include acknowledgements
% Typically, this will include thanks to reviewers who gave useful comments, to colleagues
% who contributed to the ideas, and to funding agencies and corporate sponsors that provided
% financial support.
We would like to thank the anonymous reviewers for their useful comments, and Evgenii
Egorov, Ruslan Kostoev (ADASE) and Danila Doroshin (Huawei) for fruitfull discussions.
% We thank Skoltech CDISE HPC Zhores cluster staff for providing computational infrastructure.
The authors acknowledge the use of the Skoltech CDISE HPC cluster ``Zhores'' for obtaining
the results presented in this paper.

% section* acknowledgements (end)

@@

% Why does this warrant an extra section in the appendix?
In this section we show \eqref{eq:cplx-gauss-trick}.

%notation
By $e_i\in \real \hookrightarrow \cplx$ we denote the $i$-th unit vector of dimensionality
\emph{conforming} to the matrix-vector expression it is used in, $[M]$ denotes \emph{row-major}
flattening of a matrix $M$ into a vector, i.e. in lexicographic order of its indices.
Furthermore $\diag{(\cdot)}$ embeds vectors into matrices with zeros everywhere except
the diagonal, and $\otimes$ is the Kronecker product, for which we note the following
identities $
  [P Q R] = (P \otimes R^\top) [Q]
$, $
  (P \otimes Q)^\top = (P^\top \otimes Q^\top)
$, and $
  (P \otimes R) (C \otimes S) = P Q \otimes R S
$ \citep{petersen_matrix_2012}.
% It follows from the definition of the Kronecker product
% of $P$ and $R^\top$ and the row-major order vectorization.

If we assume a factorized $\cplx$-Gaussian approximation \eqref{eq:c-gauss-vi-general}
for $
  W \in \cplx^{n\times m}
$, then $[W]$ is $\cplx$-Gaussian vector with
\begin{equation}  \label{eq:c-gauss-vi-general-vec}
  [W]
    \sim \mathcal{C}\mathcal{N}_{nm} \bigl(
      [\mu], \diag{[\Sigma]}, \diag{[C]}
    \bigr)
  \,,
\end{equation}
where with $C_{ij} = \Sigma_{ij} \xi_{ij}$, $\Sigma_{ij} \geq 0$, and $
  \lvert C_{ij} \rvert^2 \leq \Sigma_{ij}
$. Then for any $x \in \cplx^m$ and $b\in \cplx^n$ we have $
  y = W x + b
    = (I_n \otimes x^\top) [W] + b
$, whence the covariance and relation matrices of $y$ are
\begin{eqnarray}
  \bigl(I_n \otimes x^\top\bigr)
    \diag{[\Sigma]}
  \conj{\bigl(I_n \otimes x^\top\bigr)^\top}
    &=& \sum_{ij}
      \bigl(I_n \otimes x^\top\bigr)
      \biggl(
        (e_i \otimes e_j)
          \Sigma_{ij}
        (e_i \otimes e_j)^\top
      \biggr)
      \conj{\bigl(I_n \otimes x^\top\bigr)^\top}
    \notag \\
    &=& \sum_{ij}
      \bigl(e_i \otimes x^\top e_j \bigr)
        \Sigma_{ij}
      \conj{\bigl(e_i \otimes x^\top e_j \bigr)^\top}
    \notag \\
    % &=& \sum_{ij}
    %     \Sigma_{ij}
    %   \bigl(e_i e_i^\top \otimes (x^\top e_j) \conj{(x^\top e_j)^\top} \bigr)
    % \notag \\
    % &=& \sum_{i} (e_i e_i^\top) \sum_{j}
    %     \Sigma_{ij} (x^\top e_j)(e_j^\top \conj{x})
    % \notag \\
    &=&
    %   \sum_{i} (e_i e_i^\top) \sum_{j}
    %     \Sigma_{ij} (x_j \conj{x_j})
    % =
      \sum_{i=1}^n (e_i e_i^\top)
        \biggl\{
          \sum_{j=1}^m \Sigma_{ij} \lvert x_j \rvert^2
        \biggr\}
    \,,  \label{eq:cn-gauss-lrt-cov} \\
  \bigl(I_n \otimes x^\top\bigr)
    \diag{[C]}
  \bigl(I_n \otimes x^\top\bigr)^\top
    &=& \sum_{ij}
      \bigl(I_n \otimes x^\top\bigr)
      \biggl(
        (e_i \otimes e_j)
          C_{ij}
        (e_i \otimes e_j)^\top
      \biggr)
      \bigl(I_n \otimes x^\top\bigr)^\top
    \notag \\
    % &=& \sum_{ij}
    %   \bigl(e_i \otimes x^\top e_j \bigr)
    %     C_{ij}
    %   \bigl(e_i \otimes x^\top e_j \bigr)^\top
    % \notag \\
    % &=& \sum_{ij}
    %     C_{ij}
    %   \bigl(e_i e_i^\top \otimes (x^\top e_j) (x^\top e_j)^\top \bigr)
    % \notag \\
    % &=& \sum_{i} (e_i e_i^\top) \sum_{j}
    %     C_{ij} (x^\top e_j)(e_j^\top x)
    % \notag \\
    &=&
    %   \sum_{i} (e_i e_i^\top) \sum_{j}
    %     C_{ij} (x_j x_j)
    % =
      \sum_{i=1}^n (e_i e_i^\top)
        \biggl\{
          \sum_{j=1}^m C_{ij} x_j^2
        \biggr\}
    \,.  \label{eq:cn-gauss-lrt-rel}
\end{eqnarray}
Since \eqref{eq:cn-gauss-lrt-cov} and \eqref{eq:cn-gauss-lrt-rel} are diagonal, the
vector $y$ has independent univariate $\cplx$-Gaussian components, whence
\eqref{eq:cplx-gauss-trick} follows.
